---
title: "Data visualization"
author: "Jillian Allison"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# General use packages:
library(tidyverse)
library(here)
library(janitor)

# Specifically for plots:
library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)

# And for another dataset we'll explore:
library(gapminder)

# Spatial
library(sf)
```

### Read in lizard data 
```{r}
lizards <- read_csv(here("data_tidy","lizards.csv"))
```

```{r}
ggplot(data = lizards,
       aes(x = total_length, 
           y = weight)) + 
  geom_point() 
  
```

### Create a histogram of total lengths for all lizards in the data set
```{r}
ggplot(data = lizards, 
       aes(x = total_length)) + 
  geom_histogram(bins = 20, 
                 color = "orange", 
                 fill = "purple", 
                 size = 0.2, 
                 linetype = "dotted") 
# IF YOU ARENT CHANGING AN AESTHETIC MAPPED ON A VARIABLE, IT SHOULD NOT BE IN AES 
```

```{r}
ggplot(data = lizards, 
       aes(x = total_length,
           y = weight)) + 
  geom_point(
    shape = 22,
    color = "cyan4",
    fill = "yellow", 
    size = 4, 
    alpha = 0.4 # transparency 
  )
```

```{r}
ggplot(data = lizards, 
       aes(x = total_length,
           y = weight)) + 
  geom_point(aes(color = common_name, 
                 size = total_length),
             alpha = 0.5) + 
  theme_minimal() 
```

```{r}
ggplot(data = lizards, 
       aes(x = total_length,
           y = weight)) + 
  geom_point(aes(color = common_name)) + 
  facet_wrap(~common_name) + 
  theme_minimal() 
```

```{r}
lizard_total_counts_gbsum <- lizards %>% 
  group_by(common_name) %>%
  summarize(total_count = n())

lizard_total_counts_dplyr <- lizards %>%
  count(common_name)
```

```{r}
ggplot(data = lizard_total_counts_gbsum, 
       aes(y = fct_reorder(common_name, -total_count), 
           x = total_count)) + 
  geom_col() 
```

Create scatterplots, faceted by species common name, for total length versus weight. The color and shape of points should differ based on the species common name. Change the theme. Make each faceted panel have flexible axis limits. (scales should be free) 

```{r}
ggplot(data = lizards, 
       aes(x = total_length,
           y = weight)) + 
  geom_point(aes(color = common_name),
             alpha = 0.6,
             size = 1) + 
  facet_wrap(~common_name, 
             scales = "free") +
  labs(x = "Total Length (mm)", 
       y = "Weight (g)",
       title = "Jornada Basin Lizard Sizes",
       subtitle = "all lizards follow the standard length weight relationship with Western Whiptails being largest",
       caption = "Data: Jornada Basin LTER, CCBY")
  theme_minimal()
```


### Fun with scales 

```{r}
ggplot(data = lizards, 
       aes(x = total_length,
           y = weight)) + 
  geom_point(aes(color = weight)) + 
  scale_color_gradient(low = "orange", high = "purple") 

```

```{r}
ggplot(data = lizards, 
       aes(x = total_length,
           y = weight)) + 
  geom_point(aes(color = weight)) + 
  scale_color_gradientn(colors = c("magenta","cyan4","darkorchid3","dodgerblue","green","yellow")) 
```

```{r}
ggplot(data = lizards, 
       aes(x = total_length,
           y = weight)) + 
  geom_point(aes(color = weight)) + 
  scale_color_steps(low = "red", high = "black") 
```

```{r}
ggplot(data = lizards, 
       aes(x = total_length,
           y = weight)) + 
  geom_point(aes(color = total_length)) + 
  scale_color_steps2(low = "purple", 
                     mid = "white",
                     high = "orange", 
                     midpoint = 150,
                     breaks = c(50, 75, 150, 180, 220, 280))
```

```{r}
ggplot(data = lizards, 
       aes(x = total_length,
           y = weight)) + 
  geom_point(aes(color = total_length)) + 
  scale_color_stepsn(colors = c("orange", "red", "purple"), 
                     breaks = seq(from = 0, to = 300, by = 60))
```

Make a horizontal box plot with common name on the y-axis, total length on the x axis, fill changing based on the common name 

```{r}
lizards_median <- lizards %>% 
  mutate(common_name = fct_reorder(common_name, total_length, .fun = median)) # create a new column which overwrites an existing column, recasting common name as a factor. then reorder common names based on the median of total length for each of the different levels of the factor ---- common name in this case. 

ggplot(data = lizards,
       aes(x = total_length,
           y = common_name)) + 
  geom_boxplot(aes(fill = common_name))

ggplot(data = lizards_median,
       aes(y = common_name, 
           x = total_length)) + 
  geom_boxplot(aes(fill = common_name), show.legend = FALSE) + 
  scale_fill_paletteer_d(palette = "yarrr::southpark") + 
  theme_minimal() 

palette_options <- palettes_d_names
```

ugly plot
```{r}
ggplot(data = lizards,
       aes(x = total_length,
           y = weight)) + 
  geom_point(aes(color = weight)) + 
  theme(panel.grid.major.x = element_line(color = "red"),
        panel.grid.minor.y = element_blank(), 
        axis.title = element_text(color = "purple"), 
        axis.text.y = element_text(color = "green"),
        text = element_text(size = 18), 
        panel.background = element_rect(color = "purple", 
                                        fill = "green")) + 
  annotate("text", x = 300, y = 50, label = "Wooooo\ntest", color = "blue") + 
  geom_vline(xintercept = 250, linetype = "dashed", color = "orange", size = 4)
  
```


repulsive labels with `ggrepel` 
```{r}
wws_lizards <- lizards %>%
  filter(common_name == "western whiptail", 
         site == "sand") 

ggplot(data = wws_lizards,
       aes(x = total_length,
           y = weight)) + 
  geom_point() + 
  geom_text_repel(aes(label = toe_num))

gapminder %>% 
  filter(year == 2002, continent == "Europe") %>%
  ggplot(aes(x = gdpPercap, y = lifeExp)) + 
  geom_point() + 
  geom_text_repel(aes(label = country), size = 3)
```

```{r}
ggplot(data = lizards,
       aes(x = total_length,
           y = weight)) + 
  geom_point() + 
  gghighlight(weight > 30, label_key = toe_num)
```

### Make a map! 

```{r}
jornada_veg <- sf::read_sf(here("data_raw","spatial_vegetation","doc.kml"))

ggplot(data = jornada_veg) + 
  geom_sf(aes(fill = Name)) +
  theme_minimal() + 
  scale_fill_paletteer_d(palette = "ggthemes::manyeys")


```

